---
title: "Assignment 9"
author: "Shinjon Ghosh"
date: "2025-04-04"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
exer <- read.csv("https://stats.idre.ucla.edu/stat/data/exer.csv")
 ## Convert variables to factor
exer <- within(exer, {
   diet <- factor(diet)
   exertype <- factor(exertype)
    time <- factor(time)
    id <- factor(id)
})
```


```{r}
# Load Pacakage
library (ggplot2)
library(lme4)
library(nlme)
library(Matrix)
```


Question 1:
Spaghetti plot with interaction of time, diet, and exertype
```{r}
ggplot(exer, aes(x = time, y = pulse, group = id, color = diet)) +
  geom_line(alpha = 0.5) +
  facet_wrap(~exertype) +
  labs(title = "Interaction: Time x Diet x Exercise Type on Pulse") +
  theme_minimal()

```


Question 2:
Linear Model (lm)
```{r}
# Fit the linear model
lm1 <- lm(pulse ~ time + diet + exertype + time:diet + time:exertype, data = exer)

# Summary of the model
summary(lm1)

```
time2:exertype3 and time3:exertype3 are significant predictors when α = 0.05. Because p-value is less than 0.05.


Question 3:
Linear Mixed Model

```{r}
library(lme4)

lm2 <- lmer(pulse ~ time + diet + exertype*diet + exertype + (1 | id), data = exer)


summary(lm2)

```


Question 4:
Compare the fitted model to the Spaghetti plots
```{r}
exer$time_num <- as.numeric(as.character(exer$time))  # convert factor to numeric

lm3 <- lmer(pulse ~ time + diet + exertype*diet + exertype + (time_num | id), data = exer)
summary(lm3)


```

```{r}
anova(lm2, lm3)
```
The p-value is < 0.05, so adding random slopes for time significantly improves model fit.

The spaghetti plots illustrate that pulse trends over time differ not only by diet and exercise type but also by individual. In particular, individuals in Exercise Type 3 show steep and varied increases in pulse. The variability of slopes within each group suggests that modeling individual-specific slopes is appropriate.

The random slope model (lm3) fits significantly better than the random intercept model (lm2), as indicated by the likelihood ratio test (p = 0.0022). Thus, including a random slope for time captures meaningful individual differences in pulse response over time, improving model accuracy.

