---
title: "Assignment 6"
author: "Shinjon Ghosh"
date: "2025-02-21"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Libraries 
```{r}
library(MASS)
library(ggplot2)
library(randomForest)
library(xgboost)
library(lattice)
library(caret)

```


Setting seed for constant reproduction.
```{r}
set.seed(1234)
```

Question 1.
Simulate ten variables from standard normal distributions
```{r}
n <- 100
X <- as.data.frame(matrix(rnorm(n * 10), nrow = n, ncol = 10))
colnames(X) <- paste0("X", 1:10)
```

Question 2
compute the mean parameter
```{r}
mu <- 1 + 2 * X$X1 + X$X2 + 0.5 * X$X5 + 1.5 * X$X10
```

Question 3
Generate Y count response from poisson distribution
```{r}
Y <- rpois(n, lambda = exp(mu)) 
```

Combining the data into a single data frame
```{r}
data <- cbind(X, Y)
```

Question 4
Randomly splitting data into 80% training and 20% testing
```{r}
set.seed(123)
train_indices <- sample(1:n, size = 0.8 * n)
train_data <- data[train_indices, ]
test_data <- data[-train_indices, ]

# Separate predictors and response
X_train <- train_data[, -ncol(train_data)]
Y_train <- train_data$Y
X_test <- test_data[, -ncol(test_data)]
Y_test <- test_data$Y
```

Question 5
Fit Poisson regression model
```{r}
# Fit a Poisson regression model
poisson_model <- glm(Y_train ~ ., data = train_data, family = poisson())

# Check if the estimated coefficients match the actual ones
summary(poisson_model)

# Make predictions on the test dataset
poisson_predictions <- predict(poisson_model, newdata = test_data, type = "response")

```
```{r}
comparison <- data.frame(Actual = Y_test, Predicted = poisson_predictions)
print(comparison)
```


Fitting a Negative Binomial regression
```{r}
# Fit a negative binomial regression model
nb_model <- glm.nb(Y_train ~ ., data = train_data)

# Make predictions on the test dataset
nb_predictions <- predict(nb_model, newdata = test_data, type = "response")
```

Question 6 
Fit a Random forest model
```{r}

# Fit a random forest model
rf_model <- randomForest(Y_train ~ ., data = train_data)

# Make predictions on the test dataset
rf_predictions <- predict(rf_model, newdata = test_data)
```

Question 7
Fit an Extreme Gradient Boosting (XGBoost) model
```{r}
# Prepare the data for XGBoost
dtrain <- xgb.DMatrix(data = as.matrix(X_train), label = Y_train)
dtest <- xgb.DMatrix(data = as.matrix(X_test))

# Define parameters
params <- list(objective = "count:poisson", eval_metric = "rmse")

# Train the XGBoost model
xgb_model <- xgb.train(params, dtrain, nrounds = 100)

# Make predictions on the test dataset
xgb_predictions <- predict(xgb_model, dtest)
```

Question 8
Print Predictions
```{r}
print("poisson_predictions:")
print(poisson_predictions)
print("Negative Binomial Predictions:")
print(nb_predictions)
print("Random Forest Predictions:")
print(rf_predictions)
print("XGBoost Predictions:")
print(xgb_predictions)
```