---
title: "Assignment 8"
author: "Shinjon Ghosh"
date: "2025-03-28"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
exer <- read.csv("https://stats.idre.ucla.edu/stat/data/exer.csv")
 ## Convert variables to factor
exer <- within(exer, {
   diet <- factor(diet)
   exertype <- factor(exertype)
    time <- factor(time)
    id <- factor(id)
})
 
```

```{r}
# Load Pacakage
library (ggplot2)
library(lme4)
library(nlme)
library(Matrix)
```

Question 1:
```{r}
ggplot(exer, aes(x = time, y = pulse, group = id, color = diet, linetype = diet)) +
     geom_line(alpha = 0.6) +
     geom_point() +
     labs(title = "Pulse over Time by Diet Type",
          x = "Time",
          y = "Pulse",
          color = "Diet Type",
          linetype = "Diet Type") +
     theme_minimal()

```
Conclusion: Pulse tends to rise from Time 1 to Time 3, regardless of diet type.The blue dashed lines (Diet 2) show a more upward trend, especially between Time 2 and Time 3.There’s greater spread in the pulse values for Diet 2, especially at Time 3, indicating more variation in individual responses.On the other hand, Diet Type 1 shows more clustered trends.


Question 2:
```{r}
ggplot(exer, aes(x = time, y = pulse, group = id, color = exertype, linetype = exertype)) +
     geom_line(alpha = 0.6) +
    geom_point() +
     labs(title = "Pulse over Time by Exercise Type",
          x = "Time",
          y = "Pulse",
         color = "Exercise Type",
          linetype = "Exercise Type") +
     theme_minimal()

```
Conclusion:Participants in Exercise Type 3 show a sharp and consistent increase in pulse over time. Moreover, Exercise Type 2 show a mild upward trend from Time 1 to Time 3. In addition, Exercise Type 3 show minimal change in pulse over time. 


Question 3:
```{r}
# Fit linear regression model
lin <- lm(pulse ~ diet + exertype, data = exer)

# View summary of the model
summary(lin)

```
Question: 4
```{r}
# Linear regression model (fixed effects only)
 lm_model <- lm(pulse ~ diet + exertype, data = exer)
summary (lm_model)

 # Linear mixed model with random intercept for id
 lmer_model <- lmer(pulse ~ diet + exertype + (1 | id), data = exer)
 summary (lmer_model)

```

```{r}
coef(lm_model)
coef(lmer_model)

```

