---
title: "Assignment 11"
author: "Shinjon Ghosh"
date: "2025-04-18"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Libraries
```{r}
library(rstan)
library(brms)

# Simulate 100 variables from standard normal distributions
set.seed(123)
n <- 100
X <- matrix(rnorm(n * 100), nrow = n, ncol = 100)

# Calculate the mean parameter µ_i
mu <- 2 * X[, 1] + X[, 2] + 0.5 * X[, 5] + 1.5 * X[, 10]

# Generate the count response Y_i from Poisson(µ_i)
Y <- rpois(n, mu)
```


1.Stan Bayesian Model with Normal Priors
```{r}
stan_model_normal <- "
data {
  int<lower=0> N;  // number of observations
  int<lower=0> Y[N];  // count response
  matrix[N, 4] X;  // predictor matrix
}
parameters {
  vector[4] beta;  // regression coefficients
}
model {
  beta ~ normal(0, 10);  // Normal prior
  Y ~ poisson_log(X * beta);  // Poisson likelihood
}
"

# Remove any NA values from Y
valid_indices <- !is.na(Y)
X <- X[valid_indices, ]
Y <- Y[valid_indices]
n <- length(Y)

# Prepare data for Stan
stan_data <- list(N = n, Y = Y, X = X[, c(1, 2, 5, 10)])


# Fit the model
fit_normal <- stan(model_code = stan_model_normal, data = stan_data, iter = 2000, chains = 2)
print(fit_normal)
```


2. Stan Bayesian Poisson regression model with Laplace Prior
```{r}
stan_model_laplace <- "
data {
  int<lower=0> N;  // number of observations
  int<lower=0> Y[N];  // count response
  matrix[N, 4] X;  // predictor matrix
}
parameters {
  vector[4] beta;  // regression coefficients
}
model {
  beta ~ double_exponential(0, 1);  // Laplace prior
  Y ~ poisson_log(X * beta);  // Poisson likelihood
}
"

# Fit the model
fit_laplace <- stan(model_code = stan_model_laplace, data = stan_data, iter = 2000, chains = 2)
print(fit_laplace)
```


3. Comparison of coefficients in both models
```{r}
# Extract coefficients
beta_normal <- extract(fit_normal)$beta
beta_laplace <- extract(fit_laplace)$beta

# Summary statistics
summary_normal <- apply(beta_normal, 2, mean)
summary_laplace <- apply(beta_laplace, 2, mean)

# Print the comparison
comparison <- data.frame(
  Predictor = c("X1", "X2", "X5", "X10"),
  Normal_Prior = summary_normal,
  Laplace_Prior = summary_laplace
)
print(comparison)
```

From the comparison table, it can be seen that coefficient of the Normal Prior and the Laplace Prior model are approximately the same and comparing them to our specified linear equation they have significant differences in their coefficients.

